---
title: "First Markdown"
author: "Carlos"
date: "10/26/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
data(mtcars)
library(tidyverse)
library(modelr)


```
Here's a summary for mtcars. 
```{r echo=FALSE}
summary(mtcars)

```

Here's a plot of miles per gallon as a function of hp
```{r, message=FALSE, warning=FALSE, fig.align='center', fig.width=3, fig.height=2, echo=FALSE}
ggplot(mtcars, aes(x=hp,y=mpg))+
  geom_point()+
  geom_smooth(method = "lm")


```

We would like to model the effect of various factors on miles per gallon.
Let's start with a simple model where horsepower is the only predictor:

```{r,echo=FALSE}

mod1 <- glm(data = mtcars,
            formula = mpg ~ hp)


mod2 <- glm(data = mtcars,
            formula = mpg ~ poly(hp,2))

summary(mod1)
summary(mod2)


```
We built two versions of the model. The first is linear, the second is second-order
polynomial...both using hp as the only predictor.

Let's make a plot:

```{r, echo=FALSE, message=FALSE}

gather_predictions(mtcars, mod1, mod2) %>% 
  ggplot(aes(x=hp, y=mpg))+
  geom_point(size = 3, alpha = 0.25)+
  geom_point(aes(y=pred, color = model))+
  geom_smooth(aes(y=pred, color = model))

```

Which model is the best fit?
Let's look at the root mean-sq of error:

```{r}

mean(residuals(mod1)^2)
mean(residuals(mod2)^2)

```
___

Here we are fitting a curved line using natural splines with a 3rd order polynomial function...
```{r}

library(splines)
ggplot(mtcars, aes(x= hp, y=mpg))+ 
  geom_smooth(method = "lm", formula = y ~ ns(x,3))

```


[hyperlink]()




