---
title: "Assignment_9"
author: "Carlos Nunez"
date: "11/1/2020"
output: html_document
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)

#First load the libraries you will use.
library(tidyverse)
library(modelr)
library(GGally)
library(MASS)
library(broom)
library(fitdistrplus)
#Load the data.
df <- read.csv("../../Data/GradSchool_Admissions.csv", stringsAsFactors = FALSE)

```

```{r, echo=FALSE, message= FALSE, include= FALSE}

#View the data. 
head(df)

#View data structue types.
str(df)

#Clean up data
df$admit <- as.logical(df$admit) #Set the admit variable to a logical value. 

df$rank <- as.factor(df$rank) # Set the rank variable to factor. 

```

The main purpose of this analysis is to understand how significant if at all GPA, GRE, and School Rank are in affecting the chances of a student being accepted into graduate school. 

Several models were tested to find the best one which helped to show the significance that GPA, GRE, and Rank each have on the Admissions to Graduate School.The different models tested are shown below. 

```{r, echo=FALSE, message= FALSE,}

#create different logistic regression models.
model_1 <- glm(data = df, 
               formula = admit ~ gpa + gre + rank,
               family = "binomial")
model_1


model_2 <- glm(data = df, 
               formula = admit ~ gpa * gre * rank,
               family = "binomial")
model_2


model_3 <- glm(data = df,
               formula = admit ~ gpa + gre *rank,
               family = "binomial")
model_3


model_4 <- glm(data = df,
               formula = admit ~ gpa * gre + rank,
               family = "binomial")
model_4

```
P Values of each variable were also considered when analyzing the models. 

```{r,echo=FALSE, include=FALSE}
#Run summary() for each model
summary(model_1) # Has most significant values for the interaction between the exploratory variables. 
summary(model_2)
summary(model_3)
summary(model_4)

```


```{r, echo=FALSE, message= FALSE, include= FALSE}

# Run Residuals to find the best model

mean(model_1$residuals^2)# has best/ lowest value for residuals. Model one is best.  
#5.542
mean(model_2$residuals^2)
#5.921
mean(model_3$residuals^2)
#5.545
mean(model_4$residuals^2)
#5.93

```

Once the best model was determined, a prediction's function was made in order to help give more insight as to how these exploratory variables each affect the probability of being admitted into Graduate School.
Plots were created to help visualize how the exploratory variables each affected the chances of being admitted into Graduate School. These plots are shown below.  


```{r, echo = FALSE, message= FALSE}
#Once best model found make predictions
df1 <- add_predictions(df, model_1, type = "response")

#Visualize the data and look at the different relationships each exploratory variable has with admit
ggplot(df1, aes(x=gre, y=pred, color = rank))+
  geom_point()+
  theme_minimal()+
  labs(title = "Probablity of Admission vs GRE",subtitle = "Effect of GRE on Admissions",
       x ="GRE", y= "Probablity of Admission")



ggplot(df1, aes(x=gpa, y=pred, color = rank))+
  geom_point()+
  theme_minimal()+
  labs(title = "Probablity of Admission vs GPA",subtitle = "Effect of GPA on Admissions",
       x ="GPA", y= "Probablity of Admission")




ggplot(df1, aes(x=rank, y=pred, color = gpa))+
  geom_point()+
  theme_minimal()+
  labs(title = "Probablity of Admission vs School Rank",subtitle = "Effect of School Rank on Admissions",
       x ="Rank", y= "Probablity of Admission")



ggplot(df1, aes(x=rank, y=pred, color = gre))+
  geom_point()+
  theme_minimal()+
  labs(title = "Probablity of Admission vs School Rank",subtitle = "Effect of School Rank on Admissions",
       x ="Rank", y= "Probablity of Admission")

```


In conclusion all three variables are significant in determining the probability that a student will
be accepted into graduate school. Statistically from the P-values taken, and also from each respective 
plot that was made. Each plot showing the expected correlation between one of the variables and the probability of a student being accepted into Graduate School. 