---
title: "Final_Project"
author: "Carlos Nunez"
date: "12/16/2020"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 4
    code_folding: hide
---
# Apple or Microsoft?
Determining the better buy!
In this stock analysis and comparison we will be looking at two long time rivals from a value investor point of veiw. Apple and Microsoft are members of the Dow Jones, an index which represents the top 30 companies in the United States of America. For our analysis we will be using the year 2017 as a reference and also make predictions for the stock price. Also in order to help evaluate each company we will the P/E Ratio, PB Ratio, EPS(Earnings-per-share), and Current Ratio.
We will begin our analysis with Apple. First by retrieving the financial statements. We can take a peak here.
```{r setup, include=FALSE, message= FALSE}

knitr::opts_chunk$set(echo = TRUE)

#First load the libraries 
library(tidyverse)
library(tidyquant)
library(finreportr)
library(prophet)
library(patchwork)
library(jpeg)
```


```{r, echo=TRUE, include= TRUE, message=TRUE}
#Load in Financial Statements for Apple 2017
AAPL_In <- GetIncome("AAPL", 2017)

head(AAPL_In)

AAPL_Bal <- GetBalanceSheet("AAPL",2017)

head(AAPL_Bal)

```


After downloading the financial statements we will also need to obtain Apple's stock price, in order to calculate the P/E Ratio. We will also take a look at EPS. 
```{r, echo=TRUE, include=FALSE, message=FALSE}
# PE Ratio and EPS Process 

#Load Data
AAPL <- tq_get("AAPL")

head(AAPL)
#Organize data
Dates <- AAPL$date
Price <- as.numeric(AAPL$open)

#rename columns
AAPL <- data.frame(Dates = Dates , Price = Price)
head(AAPL)
#filter out years
Years <- AAPL$Dates %>% 
  as.character() %>% 
  str_split("-") %>% 
  purrr::map_chr(1)

#Add new column with just years
AAPL$Years <- Years
#filter out the year wanted
Stock <- AAPL %>% filter(Years == 2017) 


# Earnings per share process, not taking into account years where there were stock splits
Earnings <- filter(AAPL_In, Metric == "Earnings Per Share, Basic") %>% tail(2)
#pick the row that doesn't have stock split
Earnings_PS <- Earnings[1,]

```


```{r, echo=FALSE, message= FALSE, includ= FALSE}
#PB Ratio process 
Assets <- filter(AAPL_Bal, Metric == "Assets")

Liable <- filter(AAPL_Bal, Metric == "Liabilities")

Shares_out <- filter(AAPL_In, Metric == "Weighted Average Number of Shares Outstanding, Basic")
```



```{r,echo= FALSE, include=FALSE}
#Current Ratio Process
Current_As <- filter(AAPL_Bal, Metric == "Assets, Current")

Current_Li <- filter(AAPL_Bal, Metric == "Liabilities, Current")

```

Here we can see how each ratio is determined. EPS is given in the Income Statement. The P/E Ratio is given by dividing the stock price by EPS. The PB Ratio is given by substracting assets from liabilities and dividing it by the number of shares outstanding. The current ratio is by dividing current assets by the current liabilities. 
```{r, echo= TRUE, include= TRUE}
#Earnings per share 2017
Earnings_PS <- Earnings_PS$Amount

#PE Ratio 2017
PE_Ratio <- tail(Stock$Price, 1)/as.numeric(Earnings_PS)

#PB Ratio 2017
PB_Ratio <-  (as.numeric((tail(Assets$Amount, 1))) - as.numeric((tail(Liable$Amount, 1)))) / as.numeric((tail(Shares_out$Amount, 1)))

#Current Ratio 2017
Current_Ratio <- as.numeric((tail(Current_As$Amount, 1)))/ as.numeric((tail(Current_Li$Amount, 1)))

```
Now what do these ratios tell us about Apple? The P/E Ratio is a ratio used by value investor to determine if a stock is overvalued or undervalued. For a value investor a good P/E Ratio is right around the ball park of 20. It also reflects what the market currently thinks about the stock at that time. The PB ratio tells us the company's valuation relative to its book value. Book value being total assets minus intangible assets. A value investor will also look for a PB Ratio of 1 or below. The Current Ratio is used to help determine if a company has enough resources to meet its short term debts and obligations. As long as the company has a current ratio anywhere from 1.5-3 it should not have any problems meeting the short term debt and is positively looked upon by value investors. Finally EPS is representing the company's profits divided by the number of shares it has. It is a main indicator of a company's profitability. The higher, the better.  
```{r,echo= FALSE, include= FALSE, message=FALSE}
# create variables/ columns to complement data frame 
Symbol <- "AAPL"

Year <- 2017

```
 
```{r, echo= TRUE, include= TRUE, message=TRUE}
Ratios <- data.frame(Symbol, Year, Earnings_PS, PE_Ratio, PB_Ratio, Current_Ratio)

```
Here are the Financial Ratios for Apple, which we will use to compare to Microsoft. 




Here we take a peak at Microsoft Financial Statements. 
```{r, echo= TRUE, include= TRUE, message=TRUE}
#Load in Financial Statements for Microsoft 2017
MSFT_In <- GetIncome("MSFT", 2017)
head(MSFT_In)

MSFT_Bal <- GetBalanceSheet("MSFT",2017)
head(MSFT_Bal)

```

Here again we repeat the process of obtaining the values we want from the financial statements to help determine the P/E Ratio.  
```{r, echo= TRUE, include=TRUE}
#PE Ratio Process

#Load data
MSFT <- tq_get("MSFT")

head(MSFT)
#Organize data
Date <- MSFT$date
Price <- as.numeric(MSFT$open)

#Rename columns
MSFT <- data.frame(Date = Date , Price = Price)
head(MSFT)

#Filter out years
Years <- MSFT$Date %>% 
  as.character() %>% 
  str_split("-") %>% 
  purrr::map_chr(1)
#Add new column with just years
MSFT$Years <- Years
#filter out year wanted
Stock <- MSFT %>% filter(Years == 2017) 


#Earnings per share process
Earnings_PS <- filter(MSFT_In, Metric == "Earnings Per Share Basic")  %>% tail(1)

```


```{r, echo=FALSE, include= FALSE}
#PB Ratio process
Assets <- filter(MSFT_Bal, Metric == "Assets")

Liable <- filter(MSFT_Bal, Metric == "Liabilities")

Share_out <- filter(MSFT_In, Metric == "Weighted Average Number Of Shares Outstanding Basic")


```

```{r, echo=FALSE, include=FALSE}
#Current Ratio process
Current_As <- filter(MSFT_Bal, Metric == "Assets Current")

Current_Li <- filter(MSFT_Bal, Metric == "Liabilities Current")


```

Here we see how the Microsoft Financial Ratios are calculated. 
```{r, echo=TRUE, include= TRUE, message=TRUE}
#Earnings per share 2017
Earnings_PS <- Earnings_PS$Amount

#PE Ratio 2017
PE_Ratio <- tail(Stock$Price, 1)/as.numeric(Earnings_PS)

#PB Ratio 2017
PB_Ratio <-  (as.numeric((tail(Assets$Amount, 1))) - as.numeric((tail(Liable$Amount, 1)))) / as.numeric((tail(Share_out$Amount, 1)))

#Current Ratio 2017
Current_Ratio <- as.numeric((tail(Current_As$Amount, 1)))/ as.numeric((tail(Current_Li$Amount, 1)))
```

```{r, echo=FALSE, include=FALSE}
# create variables/ columns to complement data frame 
Symbol <- "MSFT"

Year <- 2017

```

```{r echo=TRUE, include=TRUE, message=TRUE}
#build data frame from ratios
Ratios1 <- data.frame(Symbol, Year, Earnings_PS, PE_Ratio, PB_Ratio, Current_Ratio)

```
Here are the Financial Ratios for Microsoft, which we will use to compare to Apple. 

## The Financial Ratios
```{r, echo=TRUE, include=TRUE, message=TRUE}
#Combine both company ratios into one data frames
Company_Ratios <- rbind(Ratios,Ratios1)

```
From these ratios we can first deduct that both companies have solid financial power. For a Value investor Apple might be a bit more appealing overall as its P/E ratio is more conservative as that from Microsoft. 



Using a software created by the FaceBook Core Data Science Team we will be making predictions of the individual companies stock price performance for the next five years. The program used is named prophet as was developed to help ordinary people make quality predictions of different events without the need of being a professional data scientist. 


Here we take a look at the predictions for Apple. 
```{r, echo=TRUE, message=FALSE, include=TRUE}
#####Stock price predictions for Apple

#Load in price data
AAPL_Pred <- tq_get("AAPL", from = "2010-01-01", to = "2017-09-30")
#view data
head(AAPL_Pred)

#Assign Variables to fit Prophet requirements
ds <- AAPL_Pred$date
y <- as.numeric(AAPL_Pred$open)

#Create new data frame that will work with Prophet
AAPL_Pred <- data.frame(ds=ds, y=y)
#view data
head(AAPL_Pred)

#Use Prophet to make stock price predictions for 5 years
mod1 <- prophet(AAPL_Pred)
future <- make_future_dataframe(mod1, period = 520)
Predictions <- predict(mod1, future)

#Plot predictions
plot(mod1,Predictions, main = "Apple Stock Price / Predictions", xlabel = "Date", ylabel = "Apple Stock Price")
#Assign to a variable
Apple_Pre <- plot(mod1,Predictions, main = "Apple Stock Price / Predictions", xlabel = "Date", ylabel = "Apple Stock Price")

```


Here we take a look at the predictions for Microsoft. 
```{r, echo=TRUE, message=FALSE, include=TRUE}
#####Stock price predictions for Microsoft

#Load in price data
MSFT_Pred <- tq_get("MSFT", from = "2010-01-01", to = "2017-09-30")
#View data
head(MSFT_Pred)

#Assign Variables to fit Prophet requirements
ds <- MSFT_Pred$date
y <- as.numeric(MSFT_Pred$open)

#Create new data frame that will work with Prophet
MSFT_Pred <- data.frame(ds=ds, y=y)
#view data
head(MSFT_Pred)

#Use Prophet to make stock price predictions for 5 years
mod2 <- prophet(MSFT_Pred)
future <- make_future_dataframe(mod2, period = 520)
Predictions1 <- predict(mod2, future)

#Plot predictions
plot(mod2, Predictions1, main="Microsoft stock price / Predictions", xlabel = "Date", ylabel = "Microsoft Stock Price")

#Assign to a variable
Microsoft_Pre <- plot(mod2, Predictions1, main="Microsoft stock price / Predictions", xlabel = "Date", ylabel = "Microsoft Stock Price")
```

## Stock Price Predictions
```{r, echo=TRUE, include=TRUE, message=TRUE}
#Both predictions side by side
Apple_Pre + Microsoft_Pre
```
As observed the Apple Company has a best case scenario stock price of $70 a share, and worst case scenario 40 dollars; While its competitor Microsoft has a best case scenario of 125 dollars per share and worst case scenario 75 dollars a share.  

# Conclusion

In conclusion from a value investor point of view Apple would be the better buy of the two rivals. Again this is mainly due to the fact that its P/E Ratio in comparison to Microsoft is much more conservative and realistic. As Microsoft could seem to be more overvalued overall. But again as mentioned previously both companies show strong financial positions. The trick for the value investor however is always to buy a wonderful company but at a 'fair' price. 
